---
title: "Bog Paper Code"
author: "Kenna Rewcastle"
date: "October 31, 2017"
output: html_document
---

```{r Read in data and remove NAs}
library(ggplot2)
library(dplyr)

setwd("~/Desktop/Classen Lab/Bog Project/Bog R") # Copy and paste into console to avoid weird Markdown issues
bog1<-read.csv("bog_2015.csv")
bog<-filter(bog1,GWC!="NA")
bogMBC<-filter(bog,MBC.mg.C.g.1.dry.soil!="NA")
```


##### Mesh size does not affect microbial biomass (plot and stats)

```{r Mesh size does not affect MBC}
meshMBC<-as.factor(bogMBC$treatment)
MBC<-bogMBC$MBC.mg.C.g.1.dry.soil

mesh_MBC<-lm(MBC~meshMBC)
anova(mesh_MBC) # p = 0.09981
summary(mesh_MBC)

mesh_cols<-c("1"="grey60","2"="grey32","3"="black")
mesh_scale<-c(1,2,3)

mesh_MBC_fig<-ggplot(data=bogMBC,aes(x=meshMBC,y=MBC,colour=meshMBC)) +
  geom_boxplot(lwd=1.5) +
  labs(x="Mesh Size",y=bquote(bold('Microbial Biomass C (mg C '*g^-1*' dry soil)'))) +
  scale_colour_manual(values=mesh_cols,labels=c("1.45 mm","55 \u03BCm","5 \u03BCm"),guide=FALSE) + 
                                                # weird text is unicode for mu
  annotate("text",x=3.4,y=4.1,label="A",size=10) +
  scale_x_discrete(limits=mesh_scale, labels=c("1.45 mm","55 \u03BCm","5 \u03BCm")) +
  theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank(),
        axis.text=element_text(colour="black",size=10),
        axis.title=element_text(size=14,face="bold"),
        panel.border=element_rect(fill=NA,colour="black",size=1.5),
        panel.background=element_rect(fill=NA))

mesh_MBC_fig
```

##### Mesh size does not affect peat water saturation (plot and stats)

```{r}
mesh<-as.factor(bog$treatment)
sat<-bog$per..saturation

mesh_sat<-lm(sat~mesh)
anova(mesh_sat) # p = 0.09372
summary(mesh_sat)

mesh_cols<-c("1"="grey60","2"="grey32","3"="black")
mesh_scale<-c(1,2,3)

mesh_sat_fig<-ggplot(data=bog,aes(x=mesh,y=sat,colour=mesh)) +
  geom_boxplot(lwd=1.5) +
  labs(x="Mesh Size",y="% Peat Water Saturation") +
  scale_colour_manual(values=mesh_cols,labels=c("1.45 mm","55 \u03BCm","5 \u03BCm"),guide=FALSE) + 
                                                # weird text is unicode for mu
  annotate("text",x=3.4,y=128,label="B",size=10) +
  scale_x_discrete(limits=mesh_scale, labels=c("1.45 mm","55 \u03BCm","5 \u03BCm")) +
  theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank(),
        axis.text=element_text(colour="black",size=10),
        axis.title=element_text(size=14,face="bold"),
        panel.border=element_rect(fill=NA,colour="black",size=1.5),
        panel.background=element_rect(fill=NA))

mesh_sat_fig
```

##### Paneled figure for supplemental Figure S1

```{r}
library(gridExtra)

grid.arrange(mesh_MBC_fig,mesh_sat_fig,nrow=1)
```

##### Compress enzymes into a total nutrient enzyme activity parameter and a total carbon enzyme activity parameter.

```{r}
# nutrient enzymes
NAG<-bogMBC$NAG
PHOS<-bogMBC$PHOS
LAP<-bogMBC$LAP 
enzyNut<-NAG+PHOS+LAP

# carbon enzymes
CBH<-bogMBC$CBH
AG<-bogMBC$AG
BG<-bogMBC$BG
enzyC<-CBH+AG+BG

# add total enzyme columns to main dataframe
realbog<-cbind(bogMBC,enzyNut,enzyC)
```

##### Labeled starch treatment was successful (increase in 13C CO2 respiration in labeled mesocosms)

```{r}
carbon<-as.factor(bog1$carbon)
d13<-bog1$D5.d13

d13effect<-t.test(formula=d13~carbon) # p < 0.001, t = -9.5717
labeled<-filter(bog1,carbon=="label")
water<-filter(bog1,carbon=="control")
mean(labeled$D5.d13) # 136.05
sd(labeled$D5.d13) # 82.06
mean(water$D5.d13) # -7.82
sd(water$D5.d13) # 6.65

### Figure for S2
d13_scale<-c(1,2)

d13effect_fig<-ggplot(data=bog1,aes(x=carbon,y=d13)) +
  geom_boxplot(lwd=1.5) +
  labs(x=expression(bold(paste(text=" "^{bold("13")},"C"," Starch Addition"))),y=expression(bold(paste("\u03B4"^{bold("13")}, "C (\u2030)")))) +
  scale_x_discrete(labels=c("control"="Control","label"="Labeled")) +
  theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank(),
        axis.text.y=element_text(colour="black",size=10),
        axis.text.x=element_text(colour="black",size=14),
        axis.title=element_text(size=14,face="bold"),
        panel.border=element_rect(fill=NA,colour="black",size=1.5),
        panel.background=element_rect(fill=NA))


```

#### Starch addition treatment doesn't affect: total resp, microbial biomass, enzyme activity (all 6), peat saturation, pH

```{r}
names(bog1)
```

```{r}
carbon<-as.factor(bog1$carbon)
resp<-bog1$D5.CO2
MBC<-bog1$MBC.mg.C.g.1.dry.soil
AG<-bog1$AG
BG<-bog1$BG
CBH<-bog1$CBH
NAG<-bog1$NAG
PHOS<-bog1$PHOS
LAP<-bog1$LAP
sat<-bog1$per..saturation
pH<-bog1$pH
DOC<-bog1$DOC..mg.C.g.1.dry.soil.
CN<-bog1$peat.C.N

# t-tests first
starch_resp<-t.test(formula=resp~carbon) # p = 0.6186, t = -0.50054
starch_MBC<-t.test(formula=MBC~carbon) # p = 0.667, t = -0.43349
starch_AG<-t.test(formula=AG~carbon) # p = 0.4964, t = 0.68441
starch_BG<-t.test(formula=BG~carbon) # p = 0.1615, t = -1.4219
starch_CBH<-t.test(formula=CBH~carbon) # p = 0.1607, t = -1.4225
starch_NAG<-t.test(formula=NAG~carbon) # p = 0.3001, t = 1.0463
starch_PHOS<-t.test(formula=PHOS~carbon) # p = 0.166, t = -1.4057
starch_LAP<-t.test(formula=LAP~carbon) # p = 0.6616, t = -0.44
starch_sat<-t.test(formula=sat~carbon) # p = 0.2097, t = -1.2694 
starch_pH<-t.test(formula=pH~carbon) # p = 0.8739, t = -0.15945
starch_DOC<-t.test(formula=DOC~carbon) # p = 0.2034, t = 1.2964
starch_CN<-t.test(formula=CN~carbon) # p = 0.9167, t = -0.10508

### MIGHT need to build a supplemental figure displaying all of this

```

##### Model selection using step AIC for resp predictors.

```{r}
resp<-realbog$D5.CO2
MBC<-realbog$MBC.mg.C.g.1.dry.soil
resp_enzyC<-realbog$enzyC
resp_enzyNut<-realbog$enzyNut
sat<-realbog$per..saturation
pH<-realbog$pH
DOC<-realbog$DOC..mg.C.g.1.dry.soil.
CN<-realbog$peat.C.N
spruce<-realbog$Dist..Spruce..m.
blub<-realbog$Dist..Blueberry..m.

# Figure out parameter order in initial model by comparing resp~paramater R2

predict1<-data.frame(MBC,resp_enzyC,resp_enzyNut,sat,pH,DOC,CN,spruce,blub)
R2vect1<-c()

param_R2<-function(predict,response,R2vect){
  N<-ncol(predict)
  for (i in 1:N) {
    regression<-lm(response~predict[,i])
    R2vect[i]<-summary(regression)$adj.r.squared
    }
  return(data.frame(names(predict),R2vect))
}

param_R2(predict=predict1,response=resp,R2vect=R2vect1)

```
```{r}
library(MASS)
resp<-realbog$D5.CO2
MBC<-realbog$MBC.mg.C.g.1.dry.soil
resp_enzyC<-realbog$enzyC
resp_enzyNut<-realbog$enzyNut
sat<-realbog$per..saturation
pH<-realbog$pH
DOC<-realbog$DOC..mg.C.g.1.dry.soil.
CN<-realbog$peat.C.N
spruce<-realbog$Dist..Spruce..m.
blub<-realbog$Dist..Blueberry..m.
  
mod001<-lm(resp~sat+ enzyNut+ MBC+ enzyC+ spruce+ DOC+ pH+ CN+ blub)
stepmod001<-stepAIC(mod001, direction="both") # best model: resp ~ sat + enzyC + spruce, AIC = 
summary(stepmod001)

```
```{r}
#### This new package computes a model selection table using AICc (better for smaller data sets)
library(AICcmodavg) 

## Take model output for each step from step AIC chart and build AICc table comparing these
Cand.mod<-list()
Cand.mod[[1]]<-lm(resp~sat+MBC+spruce+resp_PC1+DOC+CN+pH+blub)
Cand.mod[[2]]<-lm(resp~sat+MBC+spruce+resp_PC1+DOC+CN+blub)
Cand.mod[[3]]<-lm(resp~sat+MBC+spruce+resp_PC1+DOC+blub)
Cand.mod[[4]]<-lm(resp~sat+MBC+spruce+resp_PC1+blub)
Cand.mod[[5]]<-lm(resp~sat+MBC+spruce+resp_PC1)
Cand.mod[[6]]<-lm(resp~sat+spruce+resp_PC1)

## method for class 'AICclm'
aictab(cand.set=Cand.mod)

```

##### Linear regressions between parameters in best fit model and resp

```{r}
## Revert to bog1 dataset that includes NAs

resp<-bog1$D5.CO2
sat<-bog1$per..saturation
spruce<-bog1$Dist..Spruce..m.
resp_PC1<-bog_resp$resp_PC1

resp_sat<-lm(resp~sat)
summary(resp_sat) # p < 0.001, R2 = 0.40

resp_spruce<-lm(resp~spruce)
summary(resp_spruce) # p = 0.001, R2 = 0.15

resp_resp_PC1<-lm(resp~resp_PC1) # p = 0.002, R2 = 0.13
summary(resp_resp_PC1)
```

##### Model selection using step AIC for d13 predictors.

```{r}

bog_starch<-filter(realbog,carbon=="label")

d13<-bog_starch$D5.d13
MBC<-bog_starch$MBC.mg.C.g.1.dry.soil
logMBC<-log(MBC)
starch_enzyC<-bog_starch$enzyC
starch_enzyNut<-bog_starch$enzyNut
sat<-bog_starch$per..saturation
pH<-bog_starch$pH
DOC<-bog_starch$DOC..mg.C.g.1.dry.soil.
CN<-bog_starch$peat.C.N
spruce<-bog_starch$Dist..Spruce..m.
blub<-bog_starch$Dist..Blueberry..m.

# Figure out parameter order in initial model by comparing resp~paramater R2

predict2<-data.frame(logMBC,starch_enzyC,starch_enzyNut,sat,pH,DOC,CN,spruce,blub)
R2vect2<-c()


param_R2(predict=predict2,response=d13,R2vect=R2vect2)
```
```{r}
library(MASS)
  
mod002<-lm(d13~spruce+ starch_enzyC+ logMBC+ sat+ starch_enzyNut+ DOC+ CN+ pH+ blub)
stepmod002<-stepAIC(mod002, direction="both") # best model: resp~ spruce+ starch_enzyC + starch_enzyNut+ DOC
summary(stepmod002)

```
```{r}
#### This new package computes a model selection table using AICc (better for smaller data sets)
library(AICcmodavg) 

## Take model output for each step from step AIC chart and build AICc table comparing these
Cand.mod_starch<-list()
Cand.mod_starch[[1]]<-lm(d13~MBC+spruce+sat+starch_PC1+DOC+CN+pH+blub)
Cand.mod_starch[[2]]<-lm(d13~MBC+spruce+sat+starch_PC1+CN+pH+blub)
Cand.mod_starch[[3]]<-lm(d13~MBC+spruce+sat+starch_PC1+CN+pH)
Cand.mod_starch[[4]]<-lm(d13~MBC+spruce+sat+starch_PC1+pH)
Cand.mod_starch[[5]]<-lm(d13~MBC+spruce+starch_PC1+pH)
Cand.mod_starch[[6]]<-lm(d13~MBC+spruce+pH)

## method for class 'AICclm'
aictab(cand.set=Cand.mod_starch)

```

##### Linear regressions between parameters in best fit model and resp

```{r}
## Revert to bog1 dataset that includes NAs

resp<-bog1$D5.CO2
sat<-bog1$per..saturation
spruce<-bog1$Dist..Spruce..m.
resp_PC1<-bog_resp$resp_PC1

resp_sat<-lm(resp~sat)
summary(resp_sat) # p < 0.001, R2 = 0.40

resp_spruce<-lm(resp~spruce)
summary(resp_spruce) # p = 0.001, R2 = 0.15

resp_resp_PC1<-lm(resp~resp_PC1) # p = 0.002, R2 = 0.13
summary(resp_resp_PC1)
```











################################################################################################################################################### EXTRA ##################################################### #########################################################################################################

```{r}
######### JUNK ############

# From PCA help with Case
library(ggbiplot)
library(ggfortify)
library(cluster)
leaf$Group
df <- leaf[c(8,14,20,24,25)]
df
autoplot(prcomp(df), center=TRUE, scale=TRUE, obs.scale=1)
autoplot(prcomp(df), data = leaf, colour = 'Group', frame=TRUE, frame.type='norm', loadings=FALSE, loadings.label=FALSE, center=TRUE, scale=TRUE)#pca graph with   species depicted in different colors
PCA1<-prcomp(df, scale=TRUE)

Loadings <- as.data.frame(PCA1$rotation[,1:2])
summary(prcomp(df, scale=TRUE))

# Model selection
low_mod002<-lm(d13~1) # empty model to use in forward model selection
stepAIC(low_mod002,direction="forward",scope=list(upper=mod002,lower=low_mod002)) # converges on same model
```

##### Compress all enzyme activites into one PCA axis, pull out loadings for each mesocosm.

```{r}
##### This is for total respiration
names(bog1)

AG<-bog1$AG
BG<-bog1$BG
CBH<-bog1$CBH
NAG<-bog1$NAG
PHOS<-bog1$PHOS
LAP<-bog1$LAP

enzyDat<-data.frame(AG,BG,CBH,NAG,PHOS,LAP)

PCA_enzy<-prcomp(enzyDat,scale=TRUE)
enzy_loads<-data.frame(predict(PCA_enzy)) # converge to data frame so you can call out columns
resp_PC1_orig<-enzy_loads$PC1  # enzy_loads$PC1 are the PC1 loadings for each mesocosm

summary(PCA_enzy) # PC1 explains 74.45% of overall enzyme activity variation

##### This is for 13C starch

starch_bog<-filter(bog1,carbon=="label")

AG1<-starch_bog$AG
BG1<-starch_bog$BG
CBH1<-starch_bog$CBH
NAG1<-starch_bog$NAG
PHOS1<-starch_bog$PHOS
LAP1<-starch_bog$LAP

starch_enzyDat<-data.frame(AG1,BG1,CBH1,NAG1,PHOS1,LAP1)

PCA_enzy_starch<-prcomp(starch_enzyDat,scale=TRUE)
starch_enzy_loads<-data.frame(predict(PCA_enzy_starch)) # converge to data frame so you can call out columns
starch_PC1_orig<-starch_enzy_loads$PC1

summary(PCA_enzy_starch) # PC1 explains 84.31% of overall enzyme variation
```